---
output: github_document
bibliography: inst/references.bib
nocite: | 
  @brenning2012, @valavi2018
---

# mlr3spatiotempcv

Spatio-temporal resampling methods for mlr3.

<!-- badges: start -->
[![Build Status](https://img.shields.io/travis/mlr-org/mlr3spatiotempcv/master?label=Linux&logo=travis&style=flat-square)](https://travis-ci.org/mlr-org/mlr3spatiotempcv)
[![CRAN Status Badge](https://www.r-pkg.org/badges/version-ago/mlr3spatiotempcv)](https://cran.r-project.org/package=mlr3spatiotempcv)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Coverage status](https://codecov.io/gh/mlr-org/mlr3spatiotempcv/branch/master/graph/badge.svg)](https://codecov.io/github/mlr-org/mlr3spatiotempcv?branch=master)
<!-- badges: end -->

This package extends the [mlr3](https://github.com/mlr-org/mlr3) package framework by spatiotemporal resampling and visualization methods.

## Resampling methods

Currently, the following ones are implemented:

| Literature Name | R Package | Reference | mlr3spatiotempcv Name |
| --------------- | --------- | --------- | ----------------------- |
| Spatial CV  | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `ResampleSpCVCoords`  |
| Spatial Blocking | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVBlock`   |
| Environmental Blocking | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVEnv`   |
| Spatial Buffering | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVBuffer` |

## Spatial tasks

- Task "ecuador" -> `mlr_tasks$get("ecuador")`
- Task "diplodia" -> `mlr_tasks$get("diplodia")`

## Visualization methods

Generic S3 function `autoplot()` for all implemented spatial resampling methods.

The following example shows the resampling method `"spcv-coords"`. 

Visualization of all partitions

```{r setup, include=FALSE}
set.seed(123)
knitr::opts_chunk$set(fig.path = "man/figures/README-")
```


```{r spcv-coords-all-partitions}
library(mlr3)
library(mlr3spatiotempcv)
library(ggplot2)

task = tsk("ecuador")
resampling = rsmp("spcv-coords", folds = 5)
resampling$instantiate(task)

autoplot(resampling, task)
```

Visualization of the first fold only

```{r spcv-coords-fold}
autoplot(resampling, task, fold_id = 1)
```

# References
