---
output: github_document
---

# mlr3spatiotemporal

Spatio-temporal extensions for mlr3

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/mlr-org/mlr3spatiotemporal.svg?branch=master)](https://travis-ci.org/mlr-org/mlr3spatiotemporal)
[![CRAN status](https://www.r-pkg.org/badges/version/mlr3spatiotemporal)](https://cran.r-project.org/package=mlr3spatiotemporal)
[![Lifecycle: maturing](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Coverage status](https://codecov.io/gh/mlr-org/mlr3spatiotemporal/branch/master/graph/badge.svg)](https://codecov.io/github/mlr-org/mlr3spatiotemporal?branch=master)
<!-- badges: end -->

This package extends the [mlr3](https://github.com/mlr-org/mlr3) package framework by spatiotemporal resampling and visualization methods.

## Resampling methods

Currently, the following ones are implemented:


| Literature Name | R Package | Reference | mlr3spatiotemporal Name |
| --------------- | --------- | --------- | ----------------------- |
| Spatial CV  | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `ResampleSpCVCoords`  |
| Spatial Blocking | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVBlock`   |
| Environmental Blocking | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVEnv`   |
| Spatial Buffering | [blockCV](https://github.com/rvalavi/blockCV) | Valavi 2019 | `ResampleSpCVBuffer` |

## Spatial tasks

- Task "ecuador" -> `mlr_tasks$get("ecuador")`
- Task "diplodia" -> `mlr_tasks$get("diplodia")`

## Visualization methods

Generic S3 function `autoplot()` for all implemented spatial resampling methods.

The following example shows the resampling method `"spcv-coords"`. 

Visualization of all partitions

```{r setup, include=FALSE}
set.seed(123)
```


```{r spcv-coords-all-partitions}
library(mlr3)
library(mlr3spatiotemporal)
library(ggplot2)

task = tsk("ecuador")
resampling = rsmp("spcv-coords")
resampling$param_set$values = list(folds = 5)
resampling$instantiate(task)

autoplot(resampling, task)
```

Visualization of the first fold only

```{r spcv-coords-fold}
autoplot(resampling, task, fold_id = 1)
```

# References

<div id="refs" class="references">


<div id="ref-Brenning2012">

**Brenning**, A., 2012. “Spatial Cross-Validation and Bootstrap for the Assessment of
Prediction Rules in Remote Sensing: The R Package Sperrorest.” In *2012
IEEE International Geoscience and Remote Sensing Symposium*, 5372–5.
<https://doi.org/10.1109/IGARSS.2012.6352393>.

</div>

<div id="ref-Valavi2019">

**Valavi** R, Elith J, Lahoz-Monfort JJ, Guillera-Arroita G. blockCV: An r package for generating spatially or environmentally separated folds for k-fold cross-validation of species distribution models. Methods Ecol Evol. 2019; 10:225–232. doi.org/10.1111/2041-210X.13107

</div>

</div>
