---
title: "Get Started"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{foo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
set.seed(42)
library("mlr3")
library("mlr3spatiotempcv")
```

# Introduction

This package provides resampling methods for the [mlr3] package framework suited for spatial and temporal data.
**So far, only spatial resampling methods have been properly implemented.**

Calling `library("mlr3spatiotempcv")`, makes the spatial resampling methods and example tasks available to the user.

### Task types

```{r}
mlr_reflections$task_types
```

### Task column roles

```{r}
# note the addtional "coordinates" slot
mlr_reflections$task_col_roles
```

### Resampling methods

```{r, warning=FALSE}
as.data.table(mlr_resamplings)
```

Currently, the following resampling methods are implemented:

| Literature             | Package                                                   | Reference     | Class Name                      | mlr3 Sugar                     |
| ---------------        | ---------                                                 | ---------     | --------------------            | ---------------------          |
| Spatial Buffering      | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBuffer`          | `rsmp("spcv-buffer")`          |
| Spatial Blocking       | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBlock`           | `rsmp("spcv-block")`           |
| Spatial CV             | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `ResamplingSpCVCoords`          | `rsmp("spcv-coords")`          |
| Environmental Blocking | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVEnv`             | `rsmp("spcv-env")`             |
| ---                    | ---                                                       | ---           | ---                             | ---                            |
| Repeated Spatial CV    | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `RepeatedResamplingSpCVCoords` | `rsmp("repeated-spcv-coords")` |

## Usage

To create a spatial task, use class `TaskClassifST` for classification and `TaskRegrST` for regression tasks.
Please make sure that your dataset comes with coordinates since they are needed during the construction of the class (see the next code chunk).
In addition, one can specify whether the coordinates should be used as features during modeling.
<!-- coordinates are put into another object? Question is if is important for the user to know that xy-columns will be removed from the feature dataset if he/she already specified that coords should not be used as predictors.-->
By default, coordinates are not used as features and will be removed from the feature dataset after the coordinates have been created.
For optional plotting purposes, it is recommended to also pass the PROJ string specifying the used coordinate reference system via the `crs` argument.

Printing a spatial task, also displays a preview of the coordinates.

```{r}
data = mlr3::as_data_backend(readRDS(system.file("extdata", "ecuador.rda",
  package = "mlr3spatiotempcv")))

TaskClassifST$new(id = "ecuador", backend = data, target = "slides",
  positive = "TRUE", coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = "+proj=utm +zone=17 +south +datum=WGS84 +units=m +no_defs")
```

After the task has been constructed, it can be used in conjunction with all available resampling methods.
However, non-spatial methods should **only be used to showcase the degree of overoptimistic performance estimates** compared to a chosen spatiotemporal resampling method [@schratz2019].

### Resampling

To showcase the usage with code, we adapt the basic [resampling example of the mlr3book](https://mlr3book.mlr-org.com/resampling.html).

However, here we are using the "ecuador" example task that comes with the {mlr3spatiotempcv} package and use the `"repeated-spcv-coords"` resampling method.

```{r}
ecuador = tsk("ecuador")
learner = lrn("classif.rpart", maxdepth = 3, predict_type = "prob")
resampling_sp = rsmp("repeated-spcv-coords", folds = 5, repeats = 5)

rr_sp = resample(task = ecuador, learner = learner,
  resampling = resampling_sp)

rr_sp$aggregate(measures = msr("classif.ce"))
```

Now let's do the same with a classic non-spatial CV.

```{r}
resampling_nsp = rsmp("repeated_cv", folds = 5, repeats = 5)
rr_nsp = resample(task = ecuador, learner = learner,
  resampling = resampling_nsp)

rr_nsp$aggregate(measures = msr("classif.ce"))
```

We see that the classical non-spatial CV performs better.
However, "better" in this case actually means overoptimistic and biased [@burman1994; @cliff1970; @racine2000; @schratz2019].
Hence, it is advised to **not** use classical CV methods for spatiotemporal data but rather choose an appropriate spatiotemporal method.

## Visualization

This package follows the {mlr3} ecosystem idea of providing `autoplot()` methods for certain classes.

For an exhaustive overview, have a look at the [reference section for the autoplot method](https://mlr3spatiotempcv.mlr-org.com/reference/autoplot_spatial_resampling.html) on the pkgdown site of this package.

To showcase the usage, we will visualize both resampling methods which we created for the comparison we just did above.

### Spatial 

Multiple options exists.
If no specific fold is specified, all partitions will be returned distinguished by color.

```{r}
autoplot(object = resampling_sp, task = ecuador)
```

### Non-spatial

```{r, eval=FALSE}
autoplot(object = resampling_nsp, task = ecuador)
```

# References
