---
title: "Get Started"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
vignette: >
  %\VignetteIndexEntry{foo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mlr3)
library(mlr3spatiotempcv)
```

# Introduction

This package provides resampling methods for the [mlr3] package framework suited for spatial and temporal data.
**Currently, only spatial resampling are implemented.**

After calling `library("mlr3spatiotempcv")`, the additional resampling methods and example tasks of this extension package are available to the user.

### Task types

```{r}
mlr_reflections$task_types
```

### Task column roles

```{r}
# note the addtional "coordinates" slot
mlr_reflections$task_col_roles
```

### Resampling methods

```{r, warning=FALSE}
as.data.table(mlr_resamplings)
```

Currently, the following resampling methods are implemented:

| Literature             | Package                                                   | Reference     | Class Name                      | mlr3 Sugar                     |
| ---------------        | ---------                                                 | ---------     | --------------------            | ---------------------          |
| Spatial Buffering      | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBuffer`          | `rsmp("spcv-buffer")`          |
| Spatial Blocking       | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBlock`           | `rsmp("spcv-block")`           |
| Spatial CV             | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `ResamplingSpCVCoords`          | `rsmp("spcv-coords")`          |
| Environmental Blocking | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVEnv`             | `rsmp("spcv-env")`             |
| ---                    | ---                                                       | ---           | ---                             | ---                            |
| Repeated Spatial CV    | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `RepeatedRResamplingSpCVCoords` | `rsmp("repeated-spcv-coords")` |

## Usage

To create a spatial task, use the classes `TaskClassifST` and `TaskRegrST`.
Your dataset needs to come with coordinates.
The column names of these variables need to be passed during class construction via argument `coordinates_names`.

In addition, one can can specify whether the coordinates should be used as features during modelling or not.
By default they are not used and will be removed as features from the dataset after the coordinates have been created.

For optional plotting purposes, it is recommended to also pass the PROJ string via the `crs` argument to set a coordinate reference system for the supplied coordinates.

When printing a spatial task, this package adds a preview of the stored coordinates.

```{r}
data = mlr3::as_data_backend(readRDS(system.file("extdata", "ecuador.rda",
  package = "mlr3spatiotempcv")))

TaskClassifST$new(id = "ecuador", data, target = "slides",
  positive = "TRUE", coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = "+proj=utm +zone=17 +south +datum=WGS84 +units=m +no_defs")
```

After the task has been constructed, it can be used in conjunction with all resampling methods that contain `"spcv"` and all others.
However, non-spatial methods should **only be used to showcase the degree of overoptimistic performance estimates** compared to the chosen spatiotemporal resampling method [@schratz2019].



# References


