---
title: "Get Started"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{foo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
set.seed(42)
library("mlr3")
library("mlr3spatiotempcv")
```

# Introduction

This package provides resampling methods for the [mlr3] package framework suited for spatial and temporal data.
**So far, only spatial resampling has been properly implemented.**

After calling `library("mlr3spatiotempcv")`, the additional resampling methods and example tasks of this extension package are available to the user.

### Task types

```{r}
mlr_reflections$task_types
```

### Task column roles

```{r}
# note the addtional "coordinates" slot
mlr_reflections$task_col_roles
```

### Resampling methods

```{r, warning=FALSE}
as.data.table(mlr_resamplings)
```

Currently, the following resampling methods are implemented:

| Literature             | Package                                                   | Reference     | Class Name                      | mlr3 Sugar                     |
| ---------------        | ---------                                                 | ---------     | --------------------            | ---------------------          |
| Spatial Buffering      | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBuffer`          | `rsmp("spcv-buffer")`          |
| Spatial Blocking       | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVBlock`           | `rsmp("spcv-block")`           |
| Spatial CV             | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `ResamplingSpCVCoords`          | `rsmp("spcv-coords")`          |
| Environmental Blocking | [blockCV](https://github.com/rvalavi/blockCV)             | Valavi 2019   | `ResamplingSpCVEnv`             | `rsmp("spcv-env")`             |
| ---                    | ---                                                       | ---           | ---                             | ---                            |
| Repeated Spatial CV    | [sperrorest](https://github.com/giscience-fsu/sperrorest) | Brenning 2012 | `RepeatedResamplingSpCVCoords` | `rsmp("repeated-spcv-coords")` |

## Usage

To create a spatial task, use class `TaskClassifST` for classification and `TaskRegrST` for regression tasks.
Your dataset needs to come with coordinates.
The column names of these variables need to be passed during class construction via argument `coordinates_names`.

In addition, one can can specify whether the coordinates should be used as features during modeling.
<!-- coordinates are put into another R6 class?-->
By default, coordinates are not used as features and will be removed from the dataset after the coordinates have been created.

For optional plotting purposes, it is recommended to also pass the PROJ string via the `crs` argument to set a coordinate reference system for the supplied coordinates.

When printing a spatial task, this package adds a preview of the stored coordinates.

```{r}
data = mlr3::as_data_backend(readRDS(system.file("extdata", "ecuador.rda",
  package = "mlr3spatiotempcv")))

TaskClassifST$new(id = "ecuador", backend = data, target = "slides",
  positive = "TRUE", coordinate_names = c("x", "y"), coords_as_features = FALSE,
  crs = "+proj=utm +zone=17 +south +datum=WGS84 +units=m +no_defs")
```

After the task has been constructed, it can be used in conjunction with all resampling methods that contain `"spcv"` and all others.
However, non-spatial methods should **only be used to showcase the degree of overoptimistic performance estimates** compared to the chosen spatiotemporal resampling method [@schratz2019].

### Resampling

To showcase the usage with code, we adapt the basic [resampling example of the mlr3book](https://mlr3book.mlr-org.com/resampling.html).

However, here we are using the "ecuador" example task that comes with the {mlr3spatiotempcv} package and use the `"repeated-spcv-coords"` resampling method.

```{r}
ecuador = tsk("ecuador")
learner = lrn("classif.rpart", maxdepth = 3, predict_type = "prob")
resampling_sp = rsmp("repeated-spcv-coords", folds = 5, repeats = 5)

rr_sp = resample(task = ecuador, learner = learner,
  resampling = resampling_sp)

rr_sp$aggregate(measures = msr("classif.ce"))
```

Now let's do the same with a classic non-spatial CV.

```{r}
resampling_nsp = rsmp("repeated_cv", folds = 5, repeats = 5)
rr_nsp = resample(task = ecuador, learner = learner,
  resampling = resampling_nsp)

rr_nsp$aggregate(measures = msr("classif.ce"))
```

We see that the classical non-spatial CV performs better.
However, "better" in this case actually means overoptimistic and biased [@burman1994; @cliff1970; @racine2000; @schratz2019].
Hence, it is advised to **not** use classical CV methods for spatiotemporal data but rather choose an appropriate spatiotemporal method.

## Visualization

This package follows the {mlr3} ecosystem idea of providing `autoplot()` methods for certain classes.

For an exhaustive overview, have a look at the [reference section for the autoplot method](https://mlr3spatiotempcv.mlr-org.com/reference/autoplot_spatial_resampling.html) on the pkgdown site of this package.

To showcase the usage, we will visualize both resampling methods which we created for the comparison we just did above.

### Spatial 

Multiple options exists.
If no specific fold is specified, all partitions will be returned distinguished by color.

```{r}
autoplot(object = resampling_sp, task = ecuador)
```

### Non-spatial

```{r}
autoplot(object = resampling_nsp, task = ecuador)
```

# References
