---
title: "Getting Started"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
link-citations: yes
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
set.seed(42)
library("mlr3")
library("mlr3spatiotempcv")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This package adds resampling methods for the {mlr3} package framework suited for spatial, temporal and spatiotemporal data.
These methods can help to reduce the influence of [autocorrelation](https://mlr3book.mlr-org.com/spatiotemporal.html#spatiotemporal-intro) on performance estimates when performing cross-validation.
While this article gives a rather technical introduction to the package, a more applied approach can be found in the [mlr3book section on "Spatiotemporal Analysis"](https://mlr3book.mlr-org.com/spatiotemporal.html#spatiotemporal).


After loading the package via `library("mlr3spatiotempcv")`, the spatiotemporal resampling methods and example tasks provided by {mlr3spatiotempcv} are available to the user.

In mlr3, dictionaries are used for overview purposes of available methods.
The following shows which dictionaries get appended with new entries.

# Task Type

Additional task types:

-   `TaskClassifST`

-   `TaskRegrST`

```{r}
mlr_reflections$task_types
```

# Task Column Roles

Additional column roles:

-   `coordinates`

```{r}
mlr_reflections$task_col_roles
```

# Resampling Methods

Additional resampling methods:

-   `spcv_block`

-   `spcv_buffer`

-   `spcv_coords`

-   `spcv_env`

-   `sptcv_cluto`

-   `sptcv_cstf`

and their respective repeated versions.

```{r, warning=FALSE}
as.data.table(mlr_resamplings)
```

# Examples Tasks

Additional example tasks:

- `tsk("ecuador")` (spatial, classif)
- `tsk("cookfarm")` (spatiotemp, regr)

# Upstream Packages and Scientific References

The following table lists all methods implemented in {mlr3spatiotempcv}, their upstream R package and scientific references.

+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Literature                                                               | Package                                                                  | Reference                                                                | mlr3 Sugar                                                               |
+==========================================================================+==========================================================================+==========================================================================+==========================================================================+
| Spatial Buffering                                                        | [blockCV](https://github.com/rvalavi/blockCV)                            | @valavi2018                                                              | `rsmp("spcv_buffer")`                                                    |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Spatial Blocking                                                         | [blockCV](https://github.com/rvalavi/blockCV)                            | @valavi2018                                                              | `rsmp("spcv_block")`                                                     |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Spatial CV                                                               | [sperrorest](https://github.com/giscience-fsu/sperrorest)                | @brenning2012                                                            | `rsmp("spcv_coords")`                                                    |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Environmental Blocking                                                   | [blockCV](https://github.com/rvalavi/blockCV)                            | @valavi2018                                                              | `rsmp("spcv_env")`                                                       |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \-                                                                       | \-                                                                       | \-                                                                       | `rsmp("sptcv_cluto")`                                                    |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Leave-Location-and-Time-Out                                              | [CAST](https://github.com/HannaMeyer/CAST)                               | @meyer2018                                                               | `rsmp("sptcv_cstf")`                                                     |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Spatiotemporal Clustering                                                | [skmeans](https://github.com/cran/skmeans)                               | @zhao2002                                                                | `rsmp("repeated_sptcv_cluto")`                                           |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ | ------------------------------------------------------------------------ |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Repeated Spatial Blocking                                                | [blockCV](https://github.com/rvalavi/blockCV)                            | @valavi2018                                                              | `rsmp("repeated_spcv_block")`                                            |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Repeated Spatial CV                                                      | [sperrorest](https://github.com/giscience-fsu/sperrorest)                | @brenning2012                                                            | `rsmp("repeated_spcv_coords")`                                           |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Repeated Env Blocking                                                    | [blockCV](https://github.com/rvalavi/blockCV)                            | @valavi2018                                                              | `rsmp("repeated_spcv_env")`                                              |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| \-                                                                       | \-                                                                       | \-                                                                       | `rsmp("repeated_sptcv_cluto")`                                           |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Repeated Leave-Location-and-Time-Out                                     | [CAST](https://github.com/HannaMeyer/CAST)                               | @meyer2018                                                              | `rsmp("repeated_sptcv_cstf")`                                            |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+
| Repeated Spatiotemporal Clustering                                       | [skmeans](https://github.com/cran/skmeans)                               | @zhao2002                                                                | `rsmp("repeated_sptcv_cluto")`                                           |
+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+--------------------------------------------------------------------------+

# References
