---
title: "Getting Started"
output: rmarkdown::html_vignette
bibliography: ../inst/references.bib
vignette: >
  %\VignetteIndexEntry{Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
set.seed(42)
library("mlr3")
library("mlr3spatiotempcv")

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

```{r, include=FALSE}
library(mlr3spatiotempcv)
```

This package adds resampling methods for the {mlr3} package framework suited for spatial, temporal and spatiotemporal data.
These methods can help to reduce the influence of [autocorrelation](https://mlr3book.mlr-org.com/special.html#autocorrelation) on performance estimates when performing cross-validation.
While this article gives a rather technical introduction to the package, a more applied approach can be found in the [mlr3book section on "Spatiotemporal Analysis"](https://mlr3book.mlr-org.com/special.html#spatiotemporal).

After loading the package via `library("mlr3spatiotempcv")`, the spatiotemporal resampling methods and example tasks provided by {mlr3spatiotempcv} are available to the user **alongside the default {mlr3} resampling methods and tasks**.

## Creating a spatial Task

To make use of spatial resampling methods, a {mlr3} task that is aware of its spatial characteristic needs to be created.
Two `Task` child classes exist in {mlr3spatiotempcv} for this purpose:

- `TaskClassifST`
- `TaskRegrST`

To create one of these, you have multiple options:

1. Use the constructor of the `Task` directly via `$new()` - this only works for data.table backends (!)
1. Use the `as_task_*` converters (e.g. if your data is stored in an `sf` object)

We recommend the latter, as the `as_task_*` converters aim to make task construction easier, e.g., by creating the `DataBackend` (which is required to create a Task in {mlr3}) automatically and setting the `crs` and `coordinate_names` fields.
Let's assume your (point) data is stored in with an `sf` object, which is a common scenario for spatial analysis in R.

```{r 08-special-spatiotemp-002}
# create 'sf' object
data_sf = sf::st_as_sf(ecuador, coords = c("x", "y"), crs = 32717)

# create `TaskClassifST` from `sf` object
task = as_task_classif_st(data_sf, id = "ecuador_task", target = "slides", positive = "TRUE")
```

You can also use a plain `data.frame`.
In this case, `crs` and `coordinate_names` need to be passed along explicitly as they cannot be inferred directly from the `sf` object:

```{r 08-special-spatiotemp-003}
task = as_task_classif_st(ecuador, id = "ecuador_task", target = "slides",
  positive = "TRUE", coordinate_names = c("x", "y"), crs = 32717)
```

The `*ST` task family prints a subset of the coordinates by default:

```{r 08-special-spatiotemp-004}
print(task)
```

All `*ST` tasks can be treated as their super class equivalents `TaskClassif` or `TaskRegr` in subsequent {mlr3} modeling steps.

# Contributed reflections by {mlr3spatiotempcv}

In {mlr3}, dictionaries are used for overview purposes of available methods.
The following sections show which dictionaries get appended with new entries when loading {mlr3spatiotempcv}.

# Task Type

- `TaskClassifST`

- `TaskRegrST`

```{r}
mlr_reflections$task_types
```

# Task Column Roles

- `coordinate`

- `space`

- `time`

```{r}
mlr_reflections$task_col_roles
```

# Resampling Methods

* `mlr_resampling_spcv_block`

* `mlr_resampling_spcv_buffer`

* `mlr_resampling_spcv_coords`

* `mlr_resampling_spcv_disc`

* `mlr_resampling_spcv_tiles`

* `mlr_resampling_spcv_env`

* `mlr_resampling_sptcv_cstf`

and their respective repeated versions.
See `as.data.table(mlr_resamplings)` for the full dictionary.

# Examples Tasks

- `tsk("ecuador")` (spatial, classif)

- `tsk("cookfarm_mlr3")` (spatiotemp, regr)

# Upstream Packages and Scientific References

The following table lists all spatiotemporal methods implemented in {mlr3spatiotempcv} (or {mlr3}), their upstream R package and scientific references.
All methods besides `"spcv_buffer"` also have a corresponding "repeated" method.

+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| Category                       | (Package) Method Name                         | Reference         | mlr3 Notation                                                                |
+================================+===============================================+===================+==============================================================================+
| **Buffering**, spatial         | (blockCV) Spatial Buffering                   | @valavi2018       | `mlr_resamplings_spcv_buffer`                                                |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Buffering**, spatial         | (sperrorest) Spatial Disc                     | @brenning2012     | `mlr_resamplings_spcv_disc`                                                  |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Blocking**, spatial          | (blockCV) Spatial Blocking                    | @valavi2018       | `mlr_resamplings_spcv_block`                                                 |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Blocking**, spatial          | (sperrorest) Spatial Tiles                    | @valavi2018       | `mlr_resamplings_spcv_tiles`                                                 |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Clustering**, spatial        | (sperrorest) Spatial CV                       | @brenning2012     | `mlr_resamplings_spcv_coords`                                                |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Clustering**, feature-space  | (blockCV) Environmental Blocking              | @valavi2018       | `mlr_resamplings_spcv_env`                                                   |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| ------------------------------ | --------------------------------------------- | ----------------- | ---------------------------------------------------------------------------- |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Grouping**, predefined inds  | (mlr3) Predefined partitions                  |                   | `mlr_resamplings_custom_cv`                                                  |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Grouping**, spatiotemporal   | (mlr3) via `col_roles` `"group"`              |                   | `mlr_resamplings_cv`, `Task$set_col_roles(<variable>, "group")`              |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+
| **Grouping**, spatiotemporal   | (CAST) Leave-Location-and-Time-Out            | @meyer2018        | `mlr_resamplings_sptcv_cstf`, `Task$set_col_roles(<variable>, "space|time")` |
+--------------------------------+-----------------------------------------------+-------------------+------------------------------------------------------------------------------+

# References
